### 2021-03-08 paper2/3-method.tex
Lots of small improvements

\section{Interpretable Invariances by Null-Sampling}\label{ssec:general}
% \begin{figure}
%     \centering
%     \includegraphics[width=0.4\textwidth]{paper2/Figures/diagram.pdf}
%     \caption{Training procedure using the cFlow model for illustrative purposes.}%
%     \label{fig:training_diagram}
% \end{figure}
\begin{figure*}[tb]
    \centering
    \hfill
    \subfloat[cFlow model.]{%
        \scalebox{0.33}{\includegraphics[width=\textwidth]{paper2/Figures/inn_diagram_u.pdf}}%
        \label{fig:inn_diagram}
    }
    \hfill
    \subfloat[cVAE model.]{%
        \scalebox{0.4}{\includegraphics[width=\textwidth]{paper2/Figures/cvae_diagram_u.pdf}}%
        \label{fig:cvae_diagram}
    }
    \hfill
    \caption{
        Training procedure for our models. $x$: input, $s$: sensitive attribute, $z_u$: de-biased representation, $x_u$: de-biased version of the input in the data domain.
        The red bar indicates a gradient reversal layer, and $\stackrel{\rightarrow}{0}$ the null-sampling operation.
    }%
    \label{fig:model-diagrams}
\end{figure*}

\subsection{Problem Statement} % HEADING IS JUST TO HELP ORGANIZE THOUGHTS, WILL BE DELETED LATER
\noindent We assume we are given inputs $\bm{x} \in \mathcal{X}$ and corresponding labels $y \in \mathcal{Y}$.
Furthermore, there is some spurious variable $s \in \mathcal{S}$ associated with each input $\bm{x}$ which we do \emph{not} want to predict. 
Let $X$, $S$ and $Y$ be random variables that take on the values $\bm{x}$, $s$ and $y$, respectively.
The fact that both $y$ and $s$ are predictive of $\bm{x}$ implies that $\mathcal{I}(X;Y), \mathcal{I}(X;S) > 0$, where $\mathcal{I}(\cdot ;\cdot)$ is the mutual information.
Note, however, that the conditional entropy is non-zero: $H(S|X) \neq 0$, i.e., $S$ is not completely determined by $X$.

@@

We call this scenario where we only have access to the labels of a biasedly-sampled subpopulation
an \emph{aggravated fairness problem}.
These are not uncommon in the real-world.
For instance, in long-feedback systems such as mortgage-approval where the demographics of the subpopulation with observed outcomes is \emph{not} representative of the subpopulation on which the model has been deployed. 
In this case, $s$ has the potential to act as a false (or \emph{spurious}) indicator of the class label and
training a model with such a dataset would limit generalisability.
Let $(X^\mathit{tr}, S^\mathit{tr}, Y^\mathit{tr})$ then be the random variables sampled for the training set
and $(X^\mathit{te}, S^\mathit{te}, Y^\mathit{te})$ be the random variables for the test set.
The training and test sets thus induce the following inequality for their mutual information:
$\mathcal{I}(S^\mathit{tr}; Y^\mathit{tr}) \gg \mathcal{I}(S^\mathit{te}; Y^\mathit{te}) \approx 0$.

@@

\subsection{The Disentanglement Dilemma} % HEADING IS JUST TO HELP ORGANIZE THOUGHTS, WILL BE DELETED LATER
The objective in~\eqref{eq:objectivepractical} balances the two desiderata: predicting $y$ and being invariant to $s$.
However, in the training set $(X^\mathit{tr}, S^\mathit{tr}, Y^\mathit{tr})$,
$y$ and $s$ are so strongly correlated that removing information about $s$ inevitably removes information about $y$.
This strong correlation makes existing methods fail under this setting.
% However, this objective is complicated by the desideratum that $\bm{z}_u$ remain predictive of $y$,
% which precludes us from directly training on the target-labelled dataset $(X^\mathit{tr}, S^\mathit{tr}, Y^\mathit{tr})$,
%where $y$ and $s$ are so strongly correlated that removing information about $s$ inevitably removes information about $y$.
%We therefore need
In order to even define the right learning goal,
we require another source of information that allows us to disentangle $s$ and $y$.
For this, we assume the existence of another set of samples that follow a similar distribution to the test set,
but whilst the sensitive attribute is available, the class labels are not.
In reality, this is not an unreasonable assumption,
as, while properly annotated data is scarce, unlabelled data can be obtained in abundance (with demographic information from census data, electoral rolls, etc.).
Previous work has also considered treated ``unlabelled data'' as still having $s$ labels~\citep{wick2019unlocking}.
We are restricted only in the sense that the spurious correlations we want to sever are indicated in the features.
We call this the \emph{representative set}, consisting of $X^\mathit{rep}$ and $S^\mathit{rep}$.
It fulfils $\mathcal{I}(S^\mathit{rep}; Y^\mathit{rep}) \approx 0$
(or rather, it would, if the class labels $Y^\mathit{rep}$ were available).
%{\color{red} justify the existence of such a dataset}

We now summarise the training procedure; an outline for the invertible network model (cFlow) can be seen in fig.~\ref{fig:inn_diagram}.
First, the encoder network $f$ is trained on ($X^\mathit{rep}, S^\mathit{rep}$), during the first phase.
The trained network is then used to encode the training set,
taking in $\bm{x}$ and producing the representation, $\bm{z}_u$, decorrelated from the spurious variable.
The encoded dataset can then be used to train any off-the-shelf classifier safely, with information about the spurious variable having been absorbed by some auxiliary component $\mathcal{B}$.
In the case of the conditional VAE (cVAE) model,
$\mathcal{B}$ takes the form of the decoder subnetwork, which reconstructs the data conditional on a one-hot encoding of $s$,
while for the invertible network $\mathcal{B}$ is realised as a partition of the feature map $\bm{z}$
(such that $\bm{z} = [\bm{z}_u, \bm{z}_b]$), given the bijective constraint.
Thus, the classifier cannot take the shortcut of learning $s$ and instead must learn how to predict $y$ directly.
Obtaining the $s$-invariant representations, $\bm{x}_u$, in the data domain
is simply a matter of replacing the $\mathcal{B}$ component of the decoder's input for the cVAE,
and $\bm{z}_b$ for cFlow, with a zero vector of equivalent size.
We refer to this procedure used to generate $\bm{x}_u$ as \emph{null-sampling} (here, with respect to $\bm{z}_b)$.

@@

The model takes the form of a class conditional $\beta$-VAE \citep{higgins2017beta}, in which the decoder is conditioned on the spurious attribute.
We use $\theta_{enc}, \theta_{dec} \in \theta$ to denote the parameters of the encoder and decoder sub-networks, respectively.
Concretely, the encoder component performs the mapping $x \rightarrow{\bm{z}_u}$, while $\mathcal{B}$ is instantiated as the decoder,
$\mathcal{B} \coloneqq p_{\theta_{dec}}(x|z_u, s)$, which takes in a concatenation of the learned non-spurious latent vector $\bm{z}_u$ and a one-hot encoding of the spurious label $s$ to produce a reconstruction of the input $\hat{x}$.
Conditioning on a one-hot encoding of $s$, rather than a single value, as done in \citet{madras2018learning} is the key to visualising invariant representations in the data domain.
If $\mathcal{I}(z_u; s)$ is properly minimised, the decoder can only derive its information about $s$ from the label, thereby freeing up $\bm{z}_u$ from encoding the unwanted information while still allowing for reconstruction of the input.
Thus, by feeding a zero-vector to the decoder we achieve $\hat{x} \perp s$.
The full learning objective for the cVAE is given as
\begin{align}
\begin{split}
    \mathcal{L}_{\mathrm{cVAE}} =& \mathbb{E}_{q_{\theta_{enc}}(z_u, b|x)}[\log p_{\theta_{dec}}(x|z, b) - \log p_{\theta_{dec}}(s|z_u)] \\ 
    &- \beta D_{KL}(q_{\theta_{enc}}(z_u |x) \| p(z_u))
\end{split}
\end{align}
where $\beta$ is a hyperparameter that determines the trade-off between reconstruction accuracy and independence constraints,
and $p(\bm{z}_u)$ is the prior imposed on the variational posterior.
For all our experiments, $p(\bm{z}_u)$ is realised as an Isotropic Gaussian.
Fig.~\ref{fig:cvae_diagram} summarises the procedure as a diagram.

@@

The obvious solution to this is to allow the encoder to store information about $s$ in a partition of the latent space as in  \citet{creager2019flexibly}.  However, we question whether an autoencoder is the best choice for this setup, with the view that an invertible model is the better tool for the task. Using an invertible model has several guarantees, namely complete information-preservation and freedom from a reconstruction loss, the importance of which we elaborate on below.

@@

\paragraph{The Benefits of Bijectivity.}
Using an invertible network to generate our encoding, $\bm{z}_u$, carries a number of advantages over other approaches.
Ordinarily, the main benefit of flow-based models is that they permit exact density estimation. 
However, since we are not interested in sampling from the model's distribution, in our case the likelihood term serves as a regulariser, as it does for  \citet{JacSmeOya18}. 
Critically, this forces the mean of each latent dimension to zero enabling null-sampling. 
The invertible property of the network guarantees the preservation of all information relevant to $y$ which is independent of $s$, regardless of how it is allocated in the output space.
Secondly, we conjecture that the encodings are more robust to out-of-distribution data.
Whereas an autoencoder could map a previously seen input and a previously unseen input to the same representation,
an invertible network sidesteps this due to the network's bijective property, ensuring all relevant information is stored somewhere. This opens up the possibility of transfer learning between datasets with a similar manifestation of $s$, as we demonstrate in the Appendix~\ref{sec:transfer-learning}.

Under our framework, the invertible network $f$ maps the inputs $\bm{x}$ to a representation $\bm{z}_u$:
$f(\bm{x}) = \bm{z}$.
We interpret the embedding $\bm{z}$ as being the concatenation of two smaller embeddings: $\bm{z} = [\bm{z}_u, \bm{z}_b]$.
The dimensionality of $\bm{z}_b$, and $\bm{z}_u$, by complement, is a free parameter (see section~\ref{sec:optimisation-details} for tuning strategies).
As $f$ is invertible, $\bm{x}$ can be recovered like so:
\begin{align}
  \bm{x} = f^{-1}([\bm{z}_u, \bm{z}_b])
  \label{eq:zreconstruct}
\end{align}
where $\bm{z}_b$ is required for equality of the output dimension and input dimension to satisfy the bijectivity of the network -- we cannot output $\bm{z}_u$ alone, but have to output $\bm{z}_b$ as well. In order to generate the pre-image of $\bm{z}_u$, we perform null-sampling with respect to $\bm{z}_b$ by zeroing-out the elements of $\bm{z}_b$ (such that $\bm{x}_{u} = f^{-1}([\bm{z}_{u}, \stackrel{\rightarrow}{0}])$), i.e. setting them to the mean of the prior density, $\mathcal{N}(z;0, I)$.

>>>>

\section{Interpretable Invariances by Null-Sampling}\label{ssec:general}
% \begin{figure}
%     \centering
%     \includegraphics[width=0.4\textwidth]{paper2/Figures/diagram.pdf}
%     \caption{Training procedure using the cFlow model for illustrative purposes.}%
%     \label{fig:training_diagram}
% \end{figure}
\begin{figure*}[tb]
    \centering
    \hfill
    \subfloat[cFlow model.]{%
        \scalebox{0.4}{\includegraphics[width=\textwidth]{paper2/Figures/inn_diagram_u.pdf}}%
        \label{fig:inn_diagram}
    }
    \hfill
    \subfloat[cVAE model.]{%
        \scalebox{0.5}{\includegraphics[width=\textwidth]{paper2/Figures/cvae_diagram_u.pdf}}%
        \label{fig:cvae_diagram}
    }
    \hfill
    \caption{
        Training procedure for our models. $x$: input, $s$: sensitive attribute, $z_u$: de-biased representation, $x_u$: de-biased version of the input in the data domain.
        The red bar indicates a gradient reversal layer, and $\vzero$ the null-sampling operation.
    }%
    \label{fig:model-diagrams}
\end{figure*}

\subsection{Problem Statement} % HEADING IS JUST TO HELP ORGANIZE THOUGHTS, WILL BE DELETED LATER
\noindent We assume we are given inputs $\bm{x} \in \mathcal{X}$ and corresponding labels $y \in \mathcal{Y}$.
Furthermore, there is some spurious variable $s \in \mathcal{S}$ associated with each input $\bm{x}$ which we do \emph{not} want to predict. 
Let $X$, $S$ and $Y$ be random variables that take on the values $\bm{x}$, $s$ and $y$, respectively.
The fact that both $y$ and $s$ are predictive of $\bm{x}$ implies that $I(X;Y), I(X;S) > 0$, where $I(\cdot ;\cdot)$ is the mutual information.
Note, however, that the conditional entropy is non-zero: $H(S|X) \neq 0$, i.e., $S$ is not completely determined by $X$.

@@

We call this scenario where we only have access to the labels of a biasedly-sampled subpopulation
an \emph{aggravated fairness problem}.
These are not uncommon in the real-world.
For instance, in long-feedback systems such as mortgage-approval where the demographics of the subpopulation with observed outcomes is \emph{not} representative of the subpopulation on which the model has been deployed. 
In this case, $s$ has the potential to act as a false (or \emph{spurious}) indicator of the class label and
training a model with such a dataset would limit generalisability.
Let $(X^\mathit{tr}, S^\mathit{tr}, Y^\mathit{tr})$ then be the random variables sampled for the training set
and $(X^\mathit{te}, S^\mathit{te}, Y^\mathit{te})$ be the random variables for the test set.
The training and test sets thus induce the following inequality for their mutual information:
$I(S^\mathit{tr}; Y^\mathit{tr}) \gg I(S^\mathit{te}; Y^\mathit{te}) \approx 0$.

@@

\subsection{The Disentanglement Dilemma} % HEADING IS JUST TO HELP ORGANIZE THOUGHTS, WILL BE DELETED LATER
The objective in~\eqref{eq:objectivepractical} balances the two desiderata: predicting $y$ and being invariant to $s$.
However, in the training set $(X^\mathit{tr}, S^\mathit{tr}, Y^\mathit{tr})$,
$y$ and $s$ are so strongly correlated that removing information about $s$ inevitably removes information about $y$.
This strong correlation makes existing methods fail under this setting.
% However, this objective is complicated by the desideratum that $\bm{z}_u$ remain predictive of $y$,
% which precludes us from directly training on the target-labelled dataset $(X^\mathit{tr}, S^\mathit{tr}, Y^\mathit{tr})$,
%where $y$ and $s$ are so strongly correlated that removing information about $s$ inevitably removes information about $y$.
%We therefore need
In order to even define the right learning goal,
we require another source of information that allows us to disentangle $s$ and $y$.
For this, we assume the existence of another set of samples that follow a similar distribution to the test set,
but whilst the sensitive attribute is available, the class labels are not.
In reality, this is not an unreasonable assumption,
as, while properly annotated data is scarce, unlabelled data can be obtained in abundance (with demographic information from census data, electoral rolls, etc.).
Previous work has also considered treated ``unlabelled data'' as still having $s$ labels~\citep{wick2019unlocking}.
We are restricted only in the sense that the spurious correlations we want to sever are indicated in the features.
We call this the \emph{representative set}, consisting of $X^\mathit{rep}$ and $S^\mathit{rep}$.
It fulfils $I(S^\mathit{rep}; Y^\mathit{rep}) \approx 0$
(or rather, it would, if the class labels $Y^\mathit{rep}$ were available).
%{\color{red} justify the existence of such a dataset}

We now summarise the training procedure; an outline for the invertible network model (cFlow) can be seen in fig.~\ref{fig:inn_diagram}.
First, the encoder network $f$ is trained on ($X^\mathit{rep}, S^\mathit{rep}$), during the first phase.
The trained network is then used to encode the training set,
taking in $\bm{x}$ and producing the representation, $\bm{z}_u$, decorrelated from the spurious variable.
The encoded dataset can then be used to train any off-the-shelf classifier safely, with information about the spurious variable having been absorbed by some auxiliary component $\mathcal{B}$.
In the case of the \acf{cVAE} model,
$\mathcal{B}$ takes the form of the decoder subnetwork, which reconstructs the data conditional on a one-hot encoding of $s$,
while for the invertible network $\mathcal{B}$ is realised as a partition of the feature map $\bm{z}$
(such that $\bm{z} = [\bm{z}_u, \bm{z}_b]$), given the bijective constraint.
Thus, the classifier cannot take the shortcut of learning $s$ and instead must learn how to predict $y$ directly.
Obtaining the $s$-invariant representations, $\bm{x}_u$, in the data domain
is simply a matter of replacing the $\mathcal{B}$ component of the decoder's input for the cVAE,
and $\bm{z}_b$ for cFlow, with a zero vector of equivalent size.
We refer to this procedure used to generate $\bm{x}_u$ as \emph{null-sampling} (here, with respect to $\bm{z}_b)$.

@@

The model takes the form of a class conditional $\beta$-VAE \citep{higgins2017beta}, in which the decoder is conditioned on the spurious attribute.
We use $\theta_{enc}, \theta_{dec} \in \theta$ to denote the parameters of the encoder and decoder sub-networks, respectively.
Concretely, the encoder component performs the mapping $x \rightarrow{\bm{z}_u}$, while $\mathcal{B}$ is instantiated as the decoder,
$\mathcal{B} \coloneqq p_{\theta_{dec}}(x|z_u, s)$, which takes in a concatenation of the learned non-spurious latent vector $\bm{z}_u$ and a one-hot encoding of the spurious label $s$ to produce a reconstruction of the input $\hat{x}$.
Conditioning on a one-hot encoding of $s$, rather than a single value, as done in \citet{madras2018learning} is the key to visualising invariant representations in the data domain.
If $I(z_u; s)$ is properly minimised, the decoder can only derive its information about $s$ from the label, thereby freeing up $\bm{z}_u$ from encoding the unwanted information while still allowing for reconstruction of the input.
Thus, by feeding a zero-vector to the decoder we achieve $\hat{x} \perp s$.
The full learning objective for the cVAE is given as
\begin{align}
\begin{split}
    \mathcal{L}_{\mathrm{cVAE}} =& \mathbb{E}_{q_{\theta_{enc}}(z_u, b|x)}[\log p_{\theta_{dec}}(x|z, b) - \log p_{\theta_{dec}}(s|z_u)] \\ 
    &- \beta D_{KL}(q_{\theta_{enc}}(z_u |x) \| p(z_u))
\end{split}
\end{align}
where $\beta$ is a hyperparameter that determines the trade-off between reconstruction accuracy and independence constraints,
and $p(\bm{z}_u)$ is the prior imposed on the variational posterior.
For all our experiments, $p(\bm{z}_u)$ is realised as an Isotropic Gaussian.
Fig.~\ref{fig:cvae_diagram} summarises the procedure as a diagram.

@@

The obvious solution to this is to allow the encoder to store information about $s$ in a partition of the latent space as in  \citet{creager2019flexibly}.
However, we question whether an \ac{AE} is the best choice for this setup, with the view that an invertible model is the better tool for the task.
Using an invertible model has several guarantees, namely complete information-preservation and freedom from a reconstruction loss, the importance of which we elaborate on below.

@@

\paragraph{The Benefits of Bijectivity.}
Using an invertible network to generate our encoding, $\bm{z}_u$, carries a number of advantages over other approaches.
Ordinarily, the main benefit of flow-based models is that they permit exact density estimation. 
However, since we are not interested in sampling from the model's distribution, in our case the likelihood term serves as a regulariser, as it does for  \citet{JacSmeOya18}. 
Critically, this forces the mean of each latent dimension to zero enabling null-sampling. 
The invertible property of the network guarantees the preservation of all information relevant to $y$ which is independent of $s$, regardless of how it is allocated in the output space.
Secondly, we conjecture that the encodings are more robust to out-of-distribution data.
Whereas an \acf{AE} could map a previously seen input and a previously unseen input to the same representation,
an invertible network sidesteps this due to the network's bijective property, ensuring all relevant information is stored somewhere. This opens up the possibility of transfer learning between datasets with a similar manifestation of $s$, as we demonstrate in section~\ref{sec:transfer-learning}.

Under our framework, the invertible network $f$ maps the inputs $\bm{x}$ to a representation $\bm{z}_u$:
$f(\bm{x}) = \bm{z}$.
We interpret the embedding $\bm{z}$ as being the concatenation of two smaller embeddings: $\bm{z} = [\bm{z}_u, \bm{z}_b]$.
The dimensionality of $\bm{z}_b$, and $\bm{z}_u$, by complement, is a free parameter (see section~\ref{sec:optimisation-details} for tuning strategies).
As $f$ is invertible, $\bm{x}$ can be recovered like so:
\begin{align}
  \bm{x} = f^{-1}([\bm{z}_u, \bm{z}_b])
  \label{eq:zreconstruct}
\end{align}
where $\bm{z}_b$ is required for equality of the output dimension and input dimension to satisfy the bijectivity of the network -- we cannot output $\bm{z}_u$ alone, but have to output $\bm{z}_b$ as well. In order to generate the pre-image of $\bm{z}_u$, we perform null-sampling with respect to $\bm{z}_b$ by zeroing-out the elements of $\bm{z}_b$ (such that $\bm{x}_{u} = f^{-1}([\bm{z}_{u}, \vzero])$), i.e. setting them to the mean of the prior density, $\mathcal{N}(z;0, I)$.

### 2021-03-08 paper2/4-experiments.tex
Lots of small improvements

\section{Experiments}
\noindent We present experiments to demonstrate that the null-sampled representations are in fact invariant to $s$
while still allowing a classifier to predict $y$ from them.
We run our cVAE and cFlow models on the coloured MNIST (cMNIST) and CelebA dataset,
which we artificially bias, first describing the sampling procedure we follow to do so for non-synthetic datasets.
As baselines we have the model of~\citet{kim2019learning} (Ln2L) and the same CNN used to evaluate the cFlow and cVAE models
but with the unmodified images as input (CNN).
For the cFlow model we adopt a Glow-like architecture~\citep{KinDha18},
while both subnetworks of the cVAE model comprise gated convolutions~\citep{van2016conditional}, where the encoding size is $256$.
For cMNIST, we construct the Ln2L baseline according to its original description, for CelebA,
we treat it as an augmentation of the baseline CNN's objective function.
%More
Detailed information regarding model architectures can be found in Appendix~\ref{sec:architectures} and~\ref{sec:optimisation-details}.%
\footnote{The code can be found at \url{https://github.com/predictive-analytics-lab/nifr}.}
% \begin{figure*}[!htb]
%   \centering
%   %
%   \subfloat[CelebA]{\scalebox{0.495}{\includegraphics[width=\textwidth]{./results/celeba/celeba_acc_DP.pdf}}
%   \label{fig:celeba_chart}
%   }%
%   %\hfill
%   \subfloat[UCI Adult]{\scalebox{0.495}{\includegraphics[width=\textwidth]{paper2/Figures/nosinn_celeba_multiplot_all_landscape_Smiling.pdf}}
%   \label{fig:adult_chart}
%   }%
%   \caption{
%     Performance of our approach on the CelebA and UCI Adult datasets.
%     (Left of each) We show the accuracy of a downstream classifier trained on representations learned by our model.
%     (Right of each) We show the Demographic Parity of a model trained on our representations (the lower the better).
%   }\label{fig:chart}
% \end{figure*}
\begin{figure}[tb]
    \centering
    \includegraphics[width=0.7\textwidth]{paper2/Figures/nosinn_celeba.pdf}
    \caption{
        Performance of our model for different targets (mixing factor $\eta=0$).
        Left: \emph{Smiling} as target, right: \emph{high cheekbones}.
        \emph{DP diff} measures fairness with respect to demographic parity.
        A perfectly fair model has a \emph{DP diff} of 0.
    }%
    \label{fig:celeba-targets}
\end{figure}

@@

\paragraph{Results for the UCI Adult dataset.}
The UCI Adult dataset consists of census data and is commonly used to evaluate models focused on algorithmic fairness.
Following convention, we designate ``gender'' as the sensitive attribute $s$ and whether an individual's salary is \$50,000 or greater as $y$.
We show the performance of our approach in comparison to baseline approaches in fig. \ref{fig:adult-chart}.
We evaluate the performance of all models for mixing factors ($\eta$) $0$ and $1$. 
Results shown in fig. \ref{fig:adult-chart} show that
%while our model fails to surpass the baseline models in terms of accuracy for the balanced case (and those close to it),
we match or exceed the baseline.
%as $\eta$ moves the dataset towards a more imbalanced setting.
In terms of fairness metrics, our approach generally outperforms the baseline models for both of $\eta$.
Detailed results can be found in the Appendix~\ref{sec:additional-results}.

We also did experiments to show that the encoder transfers to other tasks. These transfer-learning experiments can be found in Appendix~\ref{sec:transfer-learning}.

>>>>

\section{Experiments}
\noindent We present experiments to demonstrate that the null-sampled representations are in fact invariant to $s$
while still allowing a classifier to predict $y$ from them.
We run our cVAE and cFlow models on the coloured MNIST (cMNIST) and CelebA dataset,
which we artificially bias, first describing the sampling procedure we follow to do so for non-synthetic datasets.
As baselines we have the model of~\citet{kim2019learning} (Ln2L) and the same CNN used to evaluate the cFlow and cVAE models
but with the unmodified images as input (CNN).
For the cFlow model we adopt a Glow-like architecture~\citep{KinDha18},
while both subnetworks of the cVAE model comprise gated convolutions~\citep{van2016conditional}, where the encoding size is $256$.
For cMNIST, we construct the Ln2L baseline according to its original description, for CelebA,
we treat it as an augmentation of the baseline CNN's objective function.
%More
Detailed information regarding model architectures can be found in sections~\ref{sec:architectures} and~\ref{sec:optimisation-details}.%
\footnote{The code can be found at \url{https://github.com/predictive-analytics-lab/nifr}.}
% \begin{figure*}[!htb]
%   \centering
%   %
%   \subfloat[CelebA]{\scalebox{0.495}{\includegraphics[width=\textwidth]{./results/celeba/celeba_acc_DP.pdf}}
%   \label{fig:celeba_chart}
%   }%
%   %\hfill
%   \subfloat[UCI Adult]{\scalebox{0.495}{\includegraphics[width=\textwidth]{paper2/Figures/nosinn_celeba_multiplot_all_landscape_Smiling.pdf}}
%   \label{fig:adult_chart}
%   }%
%   \caption{
%     Performance of our approach on the CelebA and UCI Adult datasets.
%     (Left of each) We show the accuracy of a downstream classifier trained on representations learned by our model.
%     (Right of each) We show the Demographic Parity of a model trained on our representations (the lower the better).
%   }\label{fig:chart}
% \end{figure*}
\begin{figure}[tb]
    \centering
    \includegraphics[width=0.7\textwidth]{paper2/Figures/nosinn_celeba.pdf}
    \caption{
        Performance of our model for different targets (mixing factor $\eta=0$).
        Left: \emph{Smiling} as target, right: \emph{high cheekbones}.
        \emph{DP diff} measures fairness with respect to demographic parity.
        A perfectly fair model has a \emph{DP diff} of 0.
    }%
    \label{fig:celeba-targets}
\end{figure}

@@

\paragraph{Results for the UCI Adult dataset.}
The UCI Adult dataset consists of census data and is commonly used to evaluate models focused on algorithmic fairness.
Following convention, we designate ``gender'' as the sensitive attribute $s$ and whether an individual's salary is \$50,000 or greater as $y$.
We show the performance of our approach in comparison to baseline approaches in fig. \ref{fig:adult-chart}.
We evaluate the performance of all models for mixing factors ($\eta$) $0$ and $1$. 
Results shown in fig. \ref{fig:adult-chart} show that
%while our model fails to surpass the baseline models in terms of accuracy for the balanced case (and those close to it),
we match or exceed the baseline.
%as $\eta$ moves the dataset towards a more imbalanced setting.
In terms of fairness metrics, our approach generally outperforms the baseline models for both of $\eta$.
Detailed results can be found in section~\ref{sec:additional-results}.

We also did experiments to show that the encoder transfers to other tasks. These transfer-learning experiments can be found in section~\ref{sec:transfer-learning}.
